R Notebook

### 2017-05-10 paper.Rnw
Add table for DIC/WAIC some minor clean ups

\noindent Bayarri, Berger, and Datta (2008) \cite{ZIP} suggest using the prior distribution $\pi(\lambda, p) \propto \frac{1}{\sqrt{\lambda}}I(0<p<1)$. This gives us the following posterior

\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:

\[
p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
\]

\noindent Neither of these distributions is recognizable. We can use a Metropolis-Hastings algorithm to sample from both of them. We will use a beta distribution as a proposal for p, and a gamma for $\lambda$. We will tune them to obtain a better acceptance rate.

The Zero Inflated Poisson (ZIP) model has two parameters. The parameter p is the probability of a structural zero, and $\lambda$ corresponds to the parameter in a typical Poisson model. For a single observation x, the probability density is:

@@

\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:

\[
p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
\]

\noindent Neither of these distributions is recognizable. We can use a Metropolis-Hastings algorithm to sample from both of them. We will use a beta distribution as a proposal for p, and a gamma for $\lambda$. We will tune them to obtain a better acceptance rate.

@@

For each variable of interest, 20,000 samples were taken. As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in. The results are as follows:

\[\]
\noindent Table 1: Results of taking 20,000 samples from the posterior distributions of $\lambda$ and p, compared between the two event types. The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses. The third does the same for p. All means and credible intervals are taken after discarding the first 10,000 samples as a burn-in.

\begin{center}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$ & p  \\ \hline
    $Tornado$ & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
    $Flash Flood$ & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
    \end{tabular}
\end{center}

\subsection{Zero Inflated Negative Binomial}
\label{ss:rZiNBinom}
<<zinbsampler, echo=FALSE>>=
@
<<densityplot, echo=FALSE>>=
@
<<zinbparams, echo=FALSE, cache=TRUE>>=
samplesize <- 650000
burnin <- 20000
thin <- 70
r.var=.00002
p.var=.00001
sigma.var=.001
@
The ZINB model was fit using the Metropolis-Hastings algorithm with a multivariate truncated normal distribution used as the proposal distribution. For the source code of the sampler, please see \ref{a:cZINB}.

This model suffered severely from the curse of dimensionality. It was very slow to converge, and the very small variance (necessary to tune the acceptance rate) led to high autocorrelation, meaning agressive thinning was necessary. Autocorrelation and running mean plots for both the Tornado and the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}. 
<<floodsample, cache=TRUE, cache.vars='flood.zinb', echo=FALSE, dependson='zinbparams'>>=
flood.zinb <- zinb.sampler(storm_details, "Flash Flood", samplesize, 
                           r.var=r.var, p.var=p.var, sigma.var=sigma.var, 
                           burnin=burnin, thinning=thin)
@
<<tornadosample, cache=TRUE, cache.vars='tornado.zinb', echo=FALSE, dependson='zinbparams'>>=
tornado.zinb <- zinb.sampler(storm_details, "Tornado", samplesize, 
                             r.var=r.var, p.var=p.var, sigma.var=sigma.var, 
                             burnin=burnin, thinning=thin)
@
In the end $\Sexpr{samplesize}$ samples were drawn originally for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, resulting in a final sample size of $\Sexpr{length(flood.zinb$sigma)}$. The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$. The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to autocorrelation and questions of convergence the results should viewed with caution.
<<evaluation, echo=FALSE, cache=FALSE>>=
@
<<model-eval, echo=FALSE, cache=TRUE, dependson=c('floodsample', 'tornadosample')>>=
flood.dic.zinb <- dic(storm_details, "Flash Flood", flood.zinb$sigma, 
                      flood.zinb$r, flood.zinb$p)
tornado.dic.zinb <- dic(storm_details, "Tornado", tornado.zinb$sigma, 
                        tornado.zinb$r, tornado.zinb$p)
#flood.wiac.zinb <- wiac(storm_details, "Flash Flood", flood.zinb$sigma, 
#                   flood.zinb$r, flood.zinb$p)
#tornado.wiac.zinb <- wiac(storm_details, "Tornado", tornado.zinb$sigma, 
#                          tornado.zinb$r, tornado.zinb$p)
@

The DIC for floods and tornadoes respectively were $\Sexpr{as.integer(flood.dic.zinb)}$ and $\Sexpr{as.integer(tornado.dic.zinb)}$. I was unable to calculate WIACs because it was too computationally expensive.
\begin{table}
\centering
\caption{The following table shows the results of the flash flood ZINB model and the tornado ZINB model. It shows the mean and the median for each parameter as well as the 95\% credible interval.}
\label{t:ZINBparams}
\begin{tabular}{lcccccc}
\toprule
&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
\cmidrule(r){2-4}\cmidrule(l){5-7}
Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
\midrule
$\sigma$ & \Sexpr{mean(flood.zinb$sigma)} &
  \Sexpr{median(flood.zinb$sigma)}&
  (\Sexpr{quantile(flood.zinb$sigma, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$sigma, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$sigma)} &
  \Sexpr{median(tornado.zinb$sigma)}&
  (\Sexpr{quantile(tornado.zinb$sigma, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$sigma, probs=c(.975))[1]})\\
$p$ &  \Sexpr{mean(flood.zinb$p)} &
  \Sexpr{median(flood.zinb$p)}&
  (\Sexpr{quantile(flood.zinb$p, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$p, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$p)} &
  \Sexpr{median(tornado.zinb$p)}&
  (\Sexpr{quantile(tornado.zinb$p, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$p, probs=c(.975))[1]})\\
$r$ &  \Sexpr{mean(flood.zinb$r)} &
  \Sexpr{median(flood.zinb$r)}&
  (\Sexpr{quantile(flood.zinb$r, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$r, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$r)} &
  \Sexpr{median(tornado.zinb$r)}&
  (\Sexpr{quantile(tornado.zinb$r, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$r, probs=c(.975))[1]})\\
\bottomrule
\end{tabular}
\end{table}
\begin{figure}
	\centering
<<floodconvergence, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=4.75, fig.width=7>>=
B<-length(flood.zinb$sigma)
Bt<-length(tornado.zinb$sigma)
par(mfrow=c(4, 3))
par(mgp=c(1.25,.5,0))
par(mar=c(3.1,4.1,2.5,.1))
plot(cumsum(flood.zinb$sigma)/1:B, 
     main = expression(paste('Cum. Mean (Flood) ',sigma)),
     ylab = 'Cumulative Mean',
     type='l')
plot(cumsum(flood.zinb$r)/1:B, 
     main = expression(paste('Cum. Mean (Flood) ',r)),
     ylab = 'Cumulative Mean',
     type='l')
plot(cumsum(flood.zinb$p)/1:B, 
     type='l', 
     main = expression(paste('Cum. Mean (Flood) ',p)),
     ylab = 'Cumulative Mean')
acf(flood.zinb$sigma, 
    main = expression(paste('ACF (Flood) ',sigma)))
acf(flood.zinb$r,
    main = expression(paste('ACF (Flood) ',r)))
acf(flood.zinb$p,
    main = expression(paste('ACF (Flood) ',p)))
plot(cumsum(tornado.zinb$sigma)/1:Bt, 
     main = expression(paste('Cum. Mean (Tornado) ',sigma)),
     ylab = 'Cumulative Mean',
     type='l')
plot(cumsum(tornado.zinb$r)/1:Bt, 
     main = expression(paste('Cum. Mean (Tornado) ',r)),
     ylab = 'Cumulative Mean',
     type='l')
plot(cumsum(tornado.zinb$p)/1:Bt, 
     type='l', 
     main = expression(paste('Cum. Mean (Tornado) ',p)),
     ylab = 'Cumulative Mean')
acf(tornado.zinb$sigma, 
    main = expression(paste('ACF (Tornado) ',sigma)))
acf(tornado.zinb$r,
    main = expression(paste('ACF (Tornado) ',r)))
acf(tornado.zinb$p,
    main = expression(paste('ACF (Tornado) ',p)))
@
	\caption{Autocorrelation plots, index plots, and cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood and the tornado data.}
	\label{f:zinbconvergence}
\end{figure}

\begin{figure}
	\centering
<<zinbdensity, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=3>>=
par(mfrow=c(1,3))
double.density.plot(flood.zinb$sigma, tornado.zinb$sigma,
                    expression(paste(sigma, ' Posterior Density ')), 
                    expression(sigma), 
                    expression(p*group("(",paste(sigma, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$p, tornado.zinb$p,
                    expression(paste(p, ' Posterior Density ')), 
                    expression(p), 
                    expression(p*group("(",paste(p, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$r, tornado.zinb$r,
                    expression(paste(r, ' Posterior Density ')), 
                    expression(r), 
                    expression(p*group("(",paste(r, "|", X),")")),
                    col=c('blue', 'black'))
@
	\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
	\label{f:zinbdensity}
\end{figure} 
\section{Discussion}
\label{s:discussion}

@@

\section{Derivations}
\label{a:derivation}
This appendix will include details on the calculations required to derive our models.
\subsection{Zero Inflated Negative Binomial Derivation}
\label{a:dZINB}
The likelihood for the ZINB is 
\begin{align*}
\mathcal{L}(X|\sigma, p, r) &= \prod_{i=1}^N \sigma I_{X=0}(X_i) + (1-\sigma)\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\\
\intertext{For ease of notation let $Z$ be the number of zero values in $X$, and $N$ be the total number of observations.}
&=\prod_{X_i=0}\left(\sigma + (1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\\
&=\left(\sigma + (1-\sigma)(1-p)^r\right)^Z\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\\
&\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
&\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^n\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)
\end{align*}
As mentioned in section \ref{ss:mZiNBinom} we take the uniform priors for $\sigma$ and $p$ as well as the non-informative gamma for $r$ which is $r^{-1/2}$. Therefore my joint posterior is:
\begin{align*}
p(r,\sigma, p|X)\propto&\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
\intertext{I am now going to take the log of the posterior because it helps with computation}
\ln\left(p(r,\sigma, p|X)\right)\propto&Z\ln\left(1 + (1/\sigma-1)(1-p)^r\right) + Z\ln(\sigma) + (N-Z)\ln(1-\sigma)+(N-Z)r\ln(1-p)\\
&+\sum_{i=1}^NX_i\ln(p)-\ln(r)/2-N\ln(\Gamma(r))+\sum_{i=1}^N\ln(\Gamma(r+X_i))
\end{align*}
\end{document}

>>>>

@@

\noindent Neither of the full conditional distributions is recognizable (see \ref{a:dZIP}). We can use a Metropolis-Hastings algorithm to sample from both of them. We will use a beta distribution as a proposal for p, and a gamma for $\lambda$. We will tune them to obtain a better acceptance rate.

@@

For each variable of interest, 20,000 samples were taken. As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in. The results are summarized in Table~\ref{t:rZIP}.

\begin{table}
    \centering
    \caption{Results of taking 20,000 samples from the posterior distributions of $\lambda$ and p, compared between the two event types. The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses. The third does the same for p. All means and credible intervals are taken after discarding the first 10,000 samples as a burn-in.}
    \label{t:rZIP}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$ & p  \\ \hline
    $Tornado$ & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
    $Flash Flood$ & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
    \end{tabular}
\end{table}

\subsection{Zero Inflated Negative Binomial}
\label{ss:rZiNBinom}
<<zinbsampler, echo=FALSE>>=
@
<<densityplot, echo=FALSE>>=
@
<<zinbparams, echo=FALSE, cache=TRUE>>=
samplesize <- 1030000
burnin <- 30000
thin <- 100
r.var=.00002
p.var=.00001
sigma.var=.001
@
The ZINB model was fit using the Metropolis-Hastings algorithm with a multivariate truncated normal distribution used as the proposal distribution. For the source code of the sampler, please see \ref{a:cZINB}.

This model suffered severely from the curse of dimensionality. It was very slow to converge, and the very small variance (necessary to tune the acceptance rate) led to high autocorrelation, meaning agressive thinning was necessary. Autocorrelation and running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}. 
<<floodsample, cache=TRUE, cache.vars='flood.zinb', echo=FALSE, dependson='zinbparams'>>=
flood.zinb <- zinb.sampler(storm_details, "Flash Flood", samplesize, 
                           r.var=r.var, p.var=p.var, sigma.var=sigma.var, 
                           burnin=burnin, thinning=thin)
@
<<tornadosample, cache=TRUE, cache.vars='tornado.zinb', echo=FALSE, dependson='zinbparams'>>=
tornado.zinb <- zinb.sampler(storm_details, "Tornado", samplesize, 
                             r.var=r.var, p.var=p.var, sigma.var=sigma.var, 
                             burnin=burnin, thinning=thin)
@
In the end $\Sexpr{samplesize}$ samples were drawn originally for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, resulting in a final sample size of $\Sexpr{length(flood.zinb$sigma)}$. The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$. The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to autocorrelation and questions of convergence the results should viewed with caution.
<<evaluation, echo=FALSE, cache=FALSE>>=
@
<<model-eval, echo=FALSE, cache=TRUE, dependson=c('floodsample', 'tornadosample')>>=
flood.dic.zinb <- dic(storm_details, "Flash Flood", flood.zinb$sigma, 
                      flood.zinb$r, flood.zinb$p)
tornado.dic.zinb <- dic(storm_details, "Tornado", tornado.zinb$sigma, 
                        tornado.zinb$r, tornado.zinb$p)
#flood.wiac.zinb <- wiac(storm_details, "Flash Flood", flood.zinb$sigma, 
#                   flood.zinb$r, flood.zinb$p)
#tornado.wiac.zinb <- wiac(storm_details, "Tornado", tornado.zinb$sigma, 
#                          tornado.zinb$r, tornado.zinb$p)
@

The DIC for floods and tornadoes respectively were $\Sexpr{as.integer(flood.dic.zinb)}$ and $\Sexpr{as.integer(tornado.dic.zinb)}$. I was unable to calculate WIACs because it was too computationally expensive.
\begin{table}
\centering
\caption{The following table shows the results of the flash flood ZINB model and the tornado ZINB model. It shows the mean and the median for each parameter as well as the 95\% credible interval.}
\label{t:ZINBparams}
\begin{tabular}{lcccccc}
\toprule
&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
\cmidrule(r){2-4}\cmidrule(l){5-7}
Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
\midrule
$\sigma$ & \Sexpr{mean(flood.zinb$sigma)} &
  \Sexpr{median(flood.zinb$sigma)}&
  (\Sexpr{quantile(flood.zinb$sigma, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$sigma, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$sigma)} &
  \Sexpr{median(tornado.zinb$sigma)}&
  (\Sexpr{quantile(tornado.zinb$sigma, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$sigma, probs=c(.975))[1]})\\
$p$ &  \Sexpr{mean(flood.zinb$p)} &
  \Sexpr{median(flood.zinb$p)}&
  (\Sexpr{quantile(flood.zinb$p, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$p, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$p)} &
  \Sexpr{median(tornado.zinb$p)}&
  (\Sexpr{quantile(tornado.zinb$p, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$p, probs=c(.975))[1]})\\
$r$ &  \Sexpr{mean(flood.zinb$r)} &
  \Sexpr{median(flood.zinb$r)}&
  (\Sexpr{quantile(flood.zinb$r, probs=c(.025))[1]}, 
    \Sexpr{quantile(flood.zinb$r, probs=c(.975))[1]})&
  \Sexpr{mean(tornado.zinb$r)} &
  \Sexpr{median(tornado.zinb$r)}&
  (\Sexpr{quantile(tornado.zinb$r, probs=c(.025))[1]}, 
    \Sexpr{quantile(tornado.zinb$r, probs=c(.975))[1]})\\
\bottomrule
\end{tabular}
\end{table}
\begin{figure}
	\centering
<<floodconvergence, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=3.5>>=
B<-length(flood.zinb$sigma)
Bt<-length(tornado.zinb$sigma)
par(mfrow=c(2, 3))
par(mgp=c(1.25,.5,0))
par(mar=c(3.1,4.1,2.5,.1))
plot(cumsum(flood.zinb$sigma)/1:B, 
     main = expression(paste('Cum. Mean (Flood) ',sigma)),
     ylab = 'Cumulative Mean',
     type='l')
plot(cumsum(flood.zinb$r)/1:B, 
     main = expression(paste('Cum. Mean (Flood) ',r)),
     ylab = 'Cumulative Mean',
     type='l')
plot(cumsum(flood.zinb$p)/1:B, 
     type='l', 
     main = expression(paste('Cum. Mean (Flood) ',p)),
     ylab = 'Cumulative Mean')
acf(flood.zinb$sigma, 
    main = expression(paste('ACF (Flood) ',sigma)))
acf(flood.zinb$r,
    main = expression(paste('ACF (Flood) ',r)))
acf(flood.zinb$p,
    main = expression(paste('ACF (Flood) ',p)))
# plot(cumsum(tornado.zinb$sigma)/1:Bt, 
#      main = expression(paste('Cum. Mean (Tornado) ',sigma)),
#      ylab = 'Cumulative Mean',
#      type='l')
# plot(cumsum(tornado.zinb$r)/1:Bt, 
#      main = expression(paste('Cum. Mean (Tornado) ',r)),
#      ylab = 'Cumulative Mean',
#      type='l')
# plot(cumsum(tornado.zinb$p)/1:Bt, 
#      type='l', 
#      main = expression(paste('Cum. Mean (Tornado) ',p)),
#      ylab = 'Cumulative Mean')
# acf(tornado.zinb$sigma, 
#     main = expression(paste('ACF (Tornado) ',sigma)))
# acf(tornado.zinb$r,
#     main = expression(paste('ACF (Tornado) ',r)))
# acf(tornado.zinb$p,
#     main = expression(paste('ACF (Tornado) ',p)))
@
	\caption{Autocorrelation plots, and cumulative mean plots for each of the three parameters in the ZINB model fit on flash flood data. The tornado plots are not pictured, but are similar in character.}
	\label{f:zinbconvergence}
\end{figure}

\begin{figure}
	\centering
<<zinbdensity, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=3>>=
par(mfrow=c(1,3))
double.density.plot(flood.zinb$sigma, tornado.zinb$sigma,
                    expression(paste(sigma, ' Posterior Density ')), 
                    expression(sigma), 
                    expression(p*group("(",paste(sigma, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$p, tornado.zinb$p,
                    expression(paste(p, ' Posterior Density ')), 
                    expression(p), 
                    expression(p*group("(",paste(p, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$r, tornado.zinb$r,
                    expression(paste(r, ' Posterior Density ')), 
                    expression(r), 
                    expression(p*group("(",paste(r, "|", X),")")),
                    col=c('blue', 'black'))
@
	\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
	\label{f:zinbdensity}
\end{figure} 
\section{Discussion}
\label{s:discussion}
The goal of this analysis was to determine which of the four models was the best fit for the data. In order to make this determination the DIC and WAIC was calculated for each model on both tornados and flash floods. The results are summarized in Table~\ref{t:evalresults}.

\begin{table}
\centering
\caption{The following table shows the resulting DIC and WAIC for each of the four models on both the tornado data and the flash flood data. In the case of the zero inflated negative binomial model, WAIC was too computationally expensive, so it was not calculated.}
\label{t:evalresults}
\begin{tabular}{lcccc}
\toprule
&\multicolumn{2}{c}{Flash Flood}&\multicolumn{2}{c}{Tornado}\\
\cmidrule(r){2-3}\cmidrule(l){4-5}
Model & DIC &WAIC & DIC &WAIC\\
\midrule
Poisson & P FLOOD DIC& P FLOOD WAIC &P TORNADO DIC&P TORNADO WAIC\\
Negative Binomial & NB FLOOD DIC& NB FLOOD WAIC &NB TORNADO DIC&NB TORNADO WAIC\\
ZIP & ZIP FLOOD DIC& ZIP FLOOD WAIC &ZIP TORNADO DIC&ZIP TORNADO WAIC\\
ZINB & \Sexpr{flood.dic.zinb}& -- &\Sexpr{tornado.dic.zinb}&--\\
\bottomrule
\end{tabular}
\end{table}

@@

\section{Derivations}
\label{a:derivation}
This appendix will include details on the calculations required to derive our models.
\subsection{Zero Inflated Poisson Derivation}
\label{a:dZIP}
\noindent In obtaining our full conditionals, we can simplify this slightly to obtain the following:

\[
p(p|X, \lambda) \propto \prod_{x_i=0}\bigg[p+(1-p)\frac{e^{-\lambda}\lambda^{x_i}}{x_i!}\bigg]\prod_{x_i \ne 0}\bigg[(1-p)\bigg]
\]
\subsection{Zero Inflated Negative Binomial Derivation}
\label{a:dZINB}
The likelihood for the ZINB is 
\begin{align*}
\mathcal{L}(X|\sigma, p, r) &= \prod_{i=1}^N \sigma I_{X=0}(X_i) + (1-\sigma)\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\\
\intertext{For ease of notation let $Z$ be the number of zero values in $X$, and $N$ be the total number of observations.}
&=\prod_{X_i=0}\left(\sigma + (1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\\
&=\left(\sigma + (1-\sigma)(1-p)^r\right)^Z\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)X_i!}\right)\\
&\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z\prod_{X_i\ne 0}\left((1-\sigma)p^{X_i}(1-p)^r\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
&\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^n\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)
\end{align*}
As mentioned in section \ref{ss:mZiNBinom} we take the uniform priors for $\sigma$ and $p$ as well as the non-informative gamma for $r$ which is $r^{-1/2}$. Therefore my joint posterior is:
\begin{align*}
p(r,\sigma, p|X)\propto&\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)\\
\intertext{I am now going to take the log of the posterior because it helps with computation}
\ln\left(p(r,\sigma, p|X)\right)\propto&Z\ln\left(1 + (1/\sigma-1)(1-p)^r\right) + Z\ln(\sigma) + (N-Z)\ln(1-\sigma)+(N-Z)r\ln(1-p)\\
&+\sum_{i=1}^NX_i\ln(p)-\ln(r)/2-N\ln(\Gamma(r))+\sum_{i=1}^N\ln(\Gamma(r+X_i))
\end{align*}
\end{document}
